---
title: "Enterovirus_B_Sequence_Curation"
output: html_document
date: "2025-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
```

Prior data collection + cleaning steps:
1) Download all Enterovirus B sequences and their metadata from NCBI Virus [All_Enterovirus_B.fasta, All_Enterovirus_B.csv]
2) Remove sequences with two or more consecutive ambiguous nucleotides (N's) in a row [All_Enterovirus_B_Curated_Sequences.fasta]
3) Identify and extract ORFs longer than 6000 nt with EMBOSS [All_Enterovirus_B_ORF_Curated_Sequences_Short_Names.fasta]
4) Codon align ORFs with RevTrans 2.0 (MAFFT) [All_Enterovirus_B_ORF_Curated_Sequences_Codon_Aligned.fasta]

#Read in data
```{r}
EVB_Info = read_csv("../All_Enterovirus_B.csv")

EVB_ORFs = readDNAStringSet("All_Enterovirus_B_ORF_Curated_Sequences_Short_Names.fasta")
EVB_Codon_Alignment = readDNAStringSet("All_Enterovirus_B_ORF_Curated_Sequences_Codon_Aligned.fasta")

#Remove extra space added at the end of each sequence name
names(EVB_Codon_Alignment) <- substr(names(EVB_Codon_Alignment), 1, nchar(names(EVB_Codon_Alignment)) - 1)
```

#Trim alignment to polyprotein ORF + remove truncated sequences
```{r}
EVB_Codon_Alignment_Trimmed = subseq(EVB_Codon_Alignment, 301, 7053)

EVB_End_Sequence = substr(EVB_Codon_Alignment_Trimmed, width(EVB_Codon_Alignment_Trimmed)[1] - 4, width(EVB_Codon_Alignment_Trimmed)[1])
Accession <- names(EVB_End_Sequence)

EVB_Truncated_Sequences = as.data.frame(cbind(Accession, EVB_End_Sequence)) %>%
  filter(EVB_End_Sequence == "-----")

`%ni%` <- Negate(`%in%`)

EVB_ORFs_Complete_Sequences = EVB_ORFs[names(EVB_ORFs) %ni% EVB_Truncated_Sequences$Accession]
writeXStringSet(EVB_ORFs_Complete_Sequences, "All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs.fasta")

EVB_Codon_Alignment_Complete_Sequences = EVB_Codon_Alignment_Trimmed[names(EVB_Codon_Alignment_Trimmed) %ni% EVB_Truncated_Sequences$Accession]
writeXStringSet(EVB_Codon_Alignment_Complete_Sequences, "All_Enterovirus_B_ORF_Curated_Sequences_Codon_Aligned_Trimmed_CompleteSeqs.fasta")
```

#Reformat Dates
```{r}
#Add dates for AY673831.1 and M16572.1 manually as the collection dates for these early sequences are not annotated in NCBI
Nancy_Strain_Dates = as.data.frame(rbind(c("AY673831.1", 1956), c("M16572.1", 1948)))
colnames(Nancy_Strain_Dates) <- c("Accession", "Collection_Date")

EVB_Dates = EVB_Info %>%
  select(Accession, Collection_Date) %>%
  rbind(Nancy_Strain_Dates) %>%
  filter(!is.na(Collection_Date)) %>%
  mutate(Accession = paste0(Accession, "_1")) %>%
  mutate(Date = ifelse(nchar(Collection_Date) == 4, paste0("7/1/", Collection_Date), 
                       ifelse(substr(Collection_Date, 5, 5) == "-", paste0(substr(Collection_Date, 6, 7), "/15/", substr(Collection_Date, 1, 4)), Collection_Date))) %>%
  mutate(Date_Format = parse_date_time(Date, "mdy")) %>%
  mutate(BEAST_Date = decimal_date(Date_Format)) %>%
  mutate(Year = trunc(BEAST_Date, 0)) %>%
  select(Accession, Year, BEAST_Date)

write_tsv(EVB_Dates, "EVB_Dates.tsv")

TreeTime_EVB_Dates = EVB_Dates %>%
  mutate(name = Accession, date = BEAST_Date) %>%
  select(name, date) 

write_csv(TreeTime_EVB_Dates, "TreeTime_EVB_Dates.csv")
```

Command line analyses:
1) Run cd-hit with c = 0.95 on All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs.fasta [All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95.fasta]

#Sequence Curation at 95% clustering
```{r}
EVB_0.95_Clusters = read_csv("Clustering_0.95/All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95.fasta.bak.clstr", col_names = c("Cluster_Length", "Accession")) %>%
  mutate(Cluster = substr(Cluster_Length, 1, nchar(Cluster_Length) - 7)) %>%
  mutate(Accession = ifelse(substr(Accession, 2, 3) == "NC", substr(Accession, 2, 14), substr(Accession, 2, 13))) %>%
  mutate(Accession = ifelse(substr(Accession, 11, 12) == "..", substr(Accession, 1, 10), Accession)) %>%
  select(Cluster, Accession)

EVB_Hosts = EVB_Info %>%
  select(Accession, Host) %>%
  mutate(Accession = paste0(Accession, "_1"))

EVB_Curation = EVB_0.95_Clusters %>%
  arrange(Accession) %>%
  left_join(EVB_Hosts, by = "Accession") %>%
  filter(Host == "Homo sapiens" | is.na(Host)) %>%
  left_join(EVB_Dates, by = "Accession") %>%
  select(Accession, Cluster, Year) %>%
  filter(!is.na(Year)) %>%
  distinct(Cluster, Year, .keep_all = TRUE) %>%
  select(Accession, Year)

write_tsv(EVB_Curation, "Clustering_0.95/Curated_EVB_Sequence_Info.tsv")

EVB_Curated_wDates = EVB_Codon_Alignment_Complete_Sequences[names(EVB_Codon_Alignment_Complete_Sequences) %in% EVB_Curation$Accession]
writeXStringSet(EVB_Curated_wDates, "Clustering_0.95/All_Enterovirus_B_ORF_Curated_Sequences_Codon_Aligned_Trimmed_CompleteSeqs_0.95.fasta")
```

Command line + online analyses:
1) Assign each EVB sequence to a type using this site: https://mpf.rivm.nl/mpf/typingtool/enterovirus/ [EVB_0.95_Types.csv]
2) Translate sequences to amino acids with EMBOSS [All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_AA.fasta]
3) Align amino acid sequences with MAFFT [All_Enterovirus_B_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_AA_Aligned.fasta]
