---
title: "Enterovirus_C_Analysis"
output: html_document
date: "2025-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggprism)
library(ggpubr)
library(ggpattern)
library(tidyquant)
library(DECIPHER)
library(QSutils)
library(coda)
```

#Read in data
```{r}
EVC_ORFs_Complete_Sequences = readDNAStringSet("All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs.fasta")

EVC_0.95_Codon_Alignment = readDNAStringSet("Clustering_0.95/All_Enterovirus_C_ORF_Curated_Sequences_Codon_Aligned_Trimmed_CompleteSeqs_0.95_woVaccine.fasta")
EVC_0.95_AA_Alignment = readAAStringSet("Clustering_0.95/All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_woVaccine_AA_Aligned.fasta")

EVC_0.95_Codon_Alignment_wVaccine = readDNAStringSet("Clustering_0.95/All_Enterovirus_C_ORF_Curated_Sequences_Codon_Aligned_Trimmed_CompleteSeqs_0.95.fasta")
  
EVC_0.95_Types = read_csv("Clustering_0.95/EVC_0.95_Types.csv") %>%
  filter(`BLAST result` == "Picornaviridae Enterovirus C") %>%
  select(name, type, `VP1 subgenogroup`)

#Write out cleaned type dataset for BEAST
write_tsv(EVC_0.95_Types, "Clustering_0.95/EVC_0.95_Types_forBEAST.tsv")

EVC_Curation_woVaccineStrains = read_tsv("Clustering_0.95/Curated_EVC_woVaccine_Sequence_Info.tsv")
```

```{r}
EVC_woVaccine_Curation = read_tsv("Clustering_0.95/Curated_EVC_woVaccine_Sequence_Info.tsv") %>%
  select(-type)

EVC_Dates = read_tsv("EVC_Dates.tsv")

EVC_woVaccine_Curation_Complete_Dates = left_join(EVC_woVaccine_Curation, EVC_Dates, join_by(Accession, Year))

write_tsv(EVC_woVaccine_Curation_Complete_Dates, "Clustering_0.95/Curated_EVC_woVaccine_Sequence_Info_Complete_Dates.tsv")
```

#Summarize types in dataset + extract type-specific sub-datasets
```{r}
EVC_0.95_Types_Summary = EVC_Curation_woVaccineStrains %>%
  group_by(type) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

#Extract accession for type with >20 representatives in the sequence set
CVA24_Accessions = EVC_0.95_Types %>%
  filter(type == "CV-A24") %>%
  select(name)

PV3_Accessions = EVC_0.95_Types %>%
  filter(type == "PV-3") %>%
  select(name)

CVA13_Accessions = EVC_0.95_Types %>%
  filter(type == "CV-A13") %>%
  select(name)

PV1_Accessions = EVC_0.95_Types %>%
  filter(type == "PV-1") %>%
  select(name)
```

#Define protein boundaries

Nucleotide
```{r}
Protein <- c("VP4", "VP2", "VP3", "VP1", "2A", "2B", "2C", "3A", "3B", "3C", "3D")
Start <- c(1, 208, 1072, 1834, 2875, 3322, 3619, 4609, 4882, 4948, 5500)
End <- c(207, 1071, 1833, 2874, 3321, 3618, 4608, 4881, 4947, 5499, 6798)
Fill <- c(0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0)

EVC_Alignment_Features = as.data.frame(cbind(Protein, Start, End, Fill)) %>%
  mutate(Start = as.numeric(Start)) %>%
  mutate(End = as.numeric(End))
```

Amino acid
```{r}
Protein <- c("VP4", "VP2", "VP3", "VP1", "2A", "2B", "2C", "3A", "3B", "3C", "3D")
Start <- c(1, 70, 344, 598, 915, 1064, 1161, 1494, 1578, 1600, 1783)
End <- c(69, 343, 597, 914, 1063, 1160, 1493, 1577, 1599, 1782, 2213)
Fill <- c(0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0)

EVC_AA_Alignment_Features = as.data.frame(cbind(Protein, Start, End, Fill)) %>%
  mutate(Start = as.numeric(Start)) %>%
  mutate(End = as.numeric(End))
```

#Entropy calculations

All Enterovirus C w/ 1 representative per type (repeated 100 times)
```{r}
EVC_1_Shannon_Entropy_Repeated = NULL

for (i in 1:100) {
  TEMP_EVC_Type_Reps = EVC_0.95_Types %>%
    sample_frac(1) %>%
    distinct(type, .keep_all = TRUE) %>%
    select(name, type) %>%
    mutate(Iteration = i)
  
  TEMP_EVC_1_AA_Alignment = EVC_0.95_AA_Alignment[names(EVC_0.95_AA_Alignment) %in% paste0(TEMP_EVC_Type_Reps$name, "_1")]
  
  TEMP_EVC_1_AA_Frequencies <- FreqMat(TEMP_EVC_1_AA_Alignment)
  Amino_Acid <- rownames(TEMP_EVC_1_AA_Frequencies)

  TEMP_EVC_1_AA_Frequencies_Gaps = cbind(Amino_Acid, as.data.frame(TEMP_EVC_1_AA_Frequencies)) %>%
    gather(key = "Site", value = "Frequency", 2:(width(TEMP_EVC_1_AA_Alignment)[1] + 1)) %>%
    mutate(Site = as.numeric(Site) - 1) %>%
    group_by(Site) %>%
    summarise(No_Gap_Sites = sum(Frequency)) %>%
    mutate(Gap_Sites = length(TEMP_EVC_1_AA_Alignment) - No_Gap_Sites) %>%
    arrange(Site)

  Gap <- TEMP_EVC_1_AA_Frequencies_Gaps$Gap_Sites

  TEMP_EVC_1_AA_Frequencies_Complete <- rbind(TEMP_EVC_1_AA_Frequencies, Gap)
  Amino_Acid_or_Gap <- rownames(TEMP_EVC_1_AA_Frequencies_Complete)
  
  TEMP_EVC_1_AA_Frequencies_DF = cbind(Amino_Acid_or_Gap, as.data.frame(TEMP_EVC_1_AA_Frequencies_Complete/length(TEMP_EVC_1_AA_Alignment))) %>%
    gather(key = "Site", value = "Frequency", 2:(width(TEMP_EVC_1_AA_Alignment)[1] + 1)) %>%
    mutate(Site = as.numeric(Site)) %>%
    filter(Frequency > 0)
  
  TEMP_EVC_1_Shannon_Entropy = TEMP_EVC_1_AA_Frequencies_DF %>% 
    mutate(Site = as.numeric(Site) - 1) %>%
    group_by(Site) %>%
    summarise(EVC_1_Entropy = -sum(Frequency * log(Frequency, base = 20))) %>%
    mutate(Iteration = i)
  
  EVC_1_Shannon_Entropy_Repeated = rbind(EVC_1_Shannon_Entropy_Repeated, TEMP_EVC_1_Shannon_Entropy)
}
```

```{r}
EVC_1_Shannon_Entropy_Summary = EVC_1_Shannon_Entropy_Repeated %>%
  group_by(Site) %>%
  summarise(Median_Entropy = median(EVC_1_Entropy), Mean_Entropy = mean(EVC_1_Entropy), SD_Entropy = sd(EVC_1_Entropy))

ggplot() +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.55, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(data = EVC_1_Shannon_Entropy_Repeated, aes(x = Site, y = EVC_1_Entropy, group = Iteration), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod", alpha = 0.1) +
  geom_ma(data = EVC_1_Shannon_Entropy_Summary, aes(x = Site, y = Mean_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  theme(legend.position = "none")
```

Coxsackievirus A24
```{r}
CVA24_AA_Accessions <- paste0(CVA24_Accessions$name, "_1")

CVA24_0.95_AA_Alignment = EVC_0.95_AA_Alignment[names(EVC_0.95_AA_Alignment) %in% CVA24_AA_Accessions]

Accessions <- names(CVA24_0.95_AA_Alignment)
CVA24_AA_Alignment_SeqNames = as.data.frame(Accessions) %>%
  mutate(Accessions = substr(Accessions, 1, nchar(Accessions) - 2))

CVA24_AA_Frequencies <- FreqMat(CVA24_0.95_AA_Alignment)
Amino_Acid <- rownames(CVA24_AA_Frequencies)

CVA24_AA_Frequencies_Gaps = cbind(Amino_Acid, as.data.frame(CVA24_AA_Frequencies)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA24_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency)) %>%
  mutate(Gap_Sites = length(CVA24_0.95_AA_Alignment) - No_Gap_Sites) %>%
  arrange(Site)

Gap <- CVA24_AA_Frequencies_Gaps$Gap_Sites

CVA24_AA_Frequencies_Complete <- rbind(CVA24_AA_Frequencies, Gap)
Amino_Acid_or_Gap <- rownames(CVA24_AA_Frequencies_Complete)

CVA24_AA_Frequencies_Check = cbind(Amino_Acid_or_Gap, as.data.frame(CVA24_AA_Frequencies_Complete)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA24_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency))
```

```{r}
CVA24_Consensus = cbind(Amino_Acid_or_Gap, as.data.frame(CVA24_AA_Frequencies_Complete/length(CVA24_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA24_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>% 
  filter(Frequency > 0) %>%
  group_by(Site) %>%
  slice_max(Frequency, n = 1) %>%
  select(Site, Amino_Acid_or_Gap) %>%
  distinct(Site, .keep_all = TRUE)

CVA24_AA_Frequencies_DF = cbind(Amino_Acid_or_Gap, as.data.frame(CVA24_AA_Frequencies_Complete/length(CVA24_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA24_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site)) %>%
  filter(Frequency > 0)
  
CVA24_Shannon_Entropy = CVA24_AA_Frequencies_DF %>% 
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(CVA24_Entropy = -sum(Frequency * log(Frequency, base = 20)), Number_of_Residues = n()) %>%
  full_join(CVA24_Consensus, by = "Site")

ggplot(CVA24_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = CVA24_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none")
```

Poliovirus 3
```{r}
PV3_AA_Accessions <- paste0(PV3_Accessions$name, "_1")

PV3_0.95_AA_Alignment = EVC_0.95_AA_Alignment[names(EVC_0.95_AA_Alignment) %in% PV3_AA_Accessions]

Accessions <- names(PV3_0.95_AA_Alignment)
PV3_AA_Alignment_SeqNames = as.data.frame(Accessions) %>%
  mutate(Accessions = substr(Accessions, 1, nchar(Accessions) - 2))

PV3_AA_Frequencies <- FreqMat(PV3_0.95_AA_Alignment)
Amino_Acid <- rownames(PV3_AA_Frequencies)

PV3_AA_Frequencies_Gaps = cbind(Amino_Acid, as.data.frame(PV3_AA_Frequencies)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV3_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency)) %>%
  mutate(Gap_Sites = length(PV3_0.95_AA_Alignment) - No_Gap_Sites) %>%
  arrange(Site)

Gap <- PV3_AA_Frequencies_Gaps$Gap_Sites

PV3_AA_Frequencies_Complete <- rbind(PV3_AA_Frequencies, Gap)
Amino_Acid_or_Gap <- rownames(PV3_AA_Frequencies_Complete)

PV3_AA_Frequencies_Check = cbind(Amino_Acid_or_Gap, as.data.frame(PV3_AA_Frequencies_Complete)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV3_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency))
```

```{r}
PV3_Consensus = cbind(Amino_Acid_or_Gap, as.data.frame(PV3_AA_Frequencies_Complete/length(PV3_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV3_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>% 
  filter(Frequency > 0) %>%
  group_by(Site) %>%
  slice_max(Frequency, n = 1) %>%
  select(Site, Amino_Acid_or_Gap) %>%
  distinct(Site, .keep_all = TRUE)

PV3_AA_Frequencies_DF = cbind(Amino_Acid_or_Gap, as.data.frame(PV3_AA_Frequencies_Complete/length(PV3_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV3_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site)) %>%
  filter(Frequency > 0)
  
PV3_Shannon_Entropy = PV3_AA_Frequencies_DF %>% 
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(PV3_Entropy = -sum(Frequency * log(Frequency, base = 20)), Number_of_Residues = n()) %>%
  full_join(PV3_Consensus, by = "Site")

ggplot(PV3_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = PV3_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none")
```

Coxsackievirus A13
```{r}
CVA13_AA_Accessions <- paste0(CVA13_Accessions$name, "_1")

CVA13_0.95_AA_Alignment = EVC_0.95_AA_Alignment[names(EVC_0.95_AA_Alignment) %in% CVA13_AA_Accessions]

Accessions <- names(CVA13_0.95_AA_Alignment)
CVA13_AA_Alignment_SeqNames = as.data.frame(Accessions) %>%
  mutate(Accessions = substr(Accessions, 1, nchar(Accessions) - 2))

CVA13_AA_Frequencies <- FreqMat(CVA13_0.95_AA_Alignment)
Amino_Acid <- rownames(CVA13_AA_Frequencies)

CVA13_AA_Frequencies_Gaps = cbind(Amino_Acid, as.data.frame(CVA13_AA_Frequencies)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA13_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency)) %>%
  mutate(Gap_Sites = length(CVA13_0.95_AA_Alignment) - No_Gap_Sites) %>%
  arrange(Site)

Gap <- CVA13_AA_Frequencies_Gaps$Gap_Sites

CVA13_AA_Frequencies_Complete <- rbind(CVA13_AA_Frequencies, Gap)
Amino_Acid_or_Gap <- rownames(CVA13_AA_Frequencies_Complete)

CVA13_AA_Frequencies_Check = cbind(Amino_Acid_or_Gap, as.data.frame(CVA13_AA_Frequencies_Complete)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA13_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency))
```

```{r}
CVA13_Consensus = cbind(Amino_Acid_or_Gap, as.data.frame(CVA13_AA_Frequencies_Complete/length(CVA13_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA13_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>% 
  filter(Frequency > 0) %>%
  group_by(Site) %>%
  slice_max(Frequency, n = 1) %>%
  select(Site, Amino_Acid_or_Gap) %>%
  distinct(Site, .keep_all = TRUE)

CVA13_AA_Frequencies_DF = cbind(Amino_Acid_or_Gap, as.data.frame(CVA13_AA_Frequencies_Complete/length(CVA13_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(CVA13_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site)) %>%
  filter(Frequency > 0)
  
CVA13_Shannon_Entropy = CVA13_AA_Frequencies_DF %>% 
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(CVA13_Entropy = -sum(Frequency * log(Frequency, base = 20)), Number_of_Residues = n()) %>%
  full_join(CVA13_Consensus, by = "Site")

ggplot(CVA13_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = CVA13_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none")
```

Poliovirus 1
```{r}
PV1_AA_Accessions <- paste0(PV1_Accessions$name, "_1")

PV1_0.95_AA_Alignment = EVC_0.95_AA_Alignment[names(EVC_0.95_AA_Alignment) %in% PV1_AA_Accessions]

Accessions <- names(PV1_0.95_AA_Alignment)
PV1_AA_Alignment_SeqNames = as.data.frame(Accessions) %>%
  mutate(Accessions = substr(Accessions, 1, nchar(Accessions) - 2))

PV1_AA_Frequencies <- FreqMat(PV1_0.95_AA_Alignment)
Amino_Acid <- rownames(PV1_AA_Frequencies)

PV1_AA_Frequencies_Gaps = cbind(Amino_Acid, as.data.frame(PV1_AA_Frequencies)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV1_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency)) %>%
  mutate(Gap_Sites = length(PV1_0.95_AA_Alignment) - No_Gap_Sites) %>%
  arrange(Site)

Gap <- PV1_AA_Frequencies_Gaps$Gap_Sites

PV1_AA_Frequencies_Complete <- rbind(PV1_AA_Frequencies, Gap)
Amino_Acid_or_Gap <- rownames(PV1_AA_Frequencies_Complete)

PV1_AA_Frequencies_Check = cbind(Amino_Acid_or_Gap, as.data.frame(PV1_AA_Frequencies_Complete)) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV1_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(No_Gap_Sites = sum(Frequency))
```

```{r}
PV1_Consensus = cbind(Amino_Acid_or_Gap, as.data.frame(PV1_AA_Frequencies_Complete/length(PV1_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV1_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site) - 1) %>% 
  filter(Frequency > 0) %>%
  group_by(Site) %>%
  slice_max(Frequency, n = 1) %>%
  select(Site, Amino_Acid_or_Gap) %>%
  distinct(Site, .keep_all = TRUE)

PV1_AA_Frequencies_DF = cbind(Amino_Acid_or_Gap, as.data.frame(PV1_AA_Frequencies_Complete/length(PV1_0.95_AA_Alignment))) %>%
  gather(key = "Site", value = "Frequency", 2:(width(PV1_0.95_AA_Alignment)[1] + 1)) %>%
  mutate(Site = as.numeric(Site)) %>%
  filter(Frequency > 0)
  
PV1_Shannon_Entropy = PV1_AA_Frequencies_DF %>% 
  mutate(Site = as.numeric(Site) - 1) %>%
  group_by(Site) %>%
  summarise(PV1_Entropy = -sum(Frequency * log(Frequency, base = 20)), Number_of_Residues = n()) %>%
  full_join(PV1_Consensus, by = "Site")

ggplot(PV1_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = PV1_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none")
```

All EVC type combined
```{r}
Position <- c(0:(width(EVC_0.95_AA_Alignment)[1] - 1))

All_EVC_Type_Shannon_Entropy = cbind(Position, CVA24_Shannon_Entropy, PV3_Shannon_Entropy, CVA13_Shannon_Entropy, PV1_Shannon_Entropy) %>%
  select(Position, CVA24_Entropy, PV3_Entropy, CVA13_Entropy, PV1_Entropy) %>%
  dplyr::rename(CVA24 = CVA24_Entropy, PV3 = PV3_Entropy, CVA13 = CVA13_Entropy, PV1 = PV1_Entropy) %>%
  gather(key = "Type", value = "Entropy", 2:5) %>%
  dplyr::rename(Site = Position) %>%
  mutate(Region = ifelse(Site <= 915, "Structural", "Nonstructural"))

pdf("All_Individual_EVC_Type_Shannon_Entropies.pdf", width = 6, height = 4.2)

ggplot(All_EVC_Type_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.175, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "lightgoldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none") +
  facet_wrap(~Type, ncol = 1, strip.position = "right")

dev.off()

All_EVC_Type_Shannon_Entropy_Region_Summary = All_EVC_Type_Shannon_Entropy %>%
  group_by(Type, Region) %>%
  summarise(Mean_Entropy = mean(Entropy)) %>%
  spread(key = Region, value = Mean_Entropy)

All_EVC_Type_Shannon_Entropy_Region_Comparison = All_EVC_Type_Shannon_Entropy %>%
  group_by(Type) %>%
  wilcox_test(Entropy ~ Region) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance("p.adj") %>%
  full_join(All_EVC_Type_Shannon_Entropy_Region_Summary, by = "Type") %>%
  select(Type, Structural, Nonstructural, p.adj, p.adj.signif) 

write_csv(All_EVC_Type_Shannon_Entropy_Region_Comparison, "All_EVC_Type_Shannon_Entropy_Region_Comparison.csv")

EVC_Type_Shannon_Entropy = All_EVC_Type_Shannon_Entropy %>%
  group_by(Site) %>%
  summarise(Average_Type_Entropy = mean(Entropy), SE_Type_Entropy = sd(Entropy)/sqrt(n()))

ggplot(EVC_Type_Shannon_Entropy) +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(aes(x = Site, y = Average_Type_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "lightgoldenrod") +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  ylab("") +
  theme(legend.position = "none")
```

By Type vs. All EVC Sequence per Type Entropy Comparison
```{r}
Type_EVC_1_Entropy_Comparison = EVC_Type_Shannon_Entropy %>%
  left_join(EVC_1_Shannon_Entropy_Summary, by = "Site") %>%
  select(Site, Average_Type_Entropy, Mean_Entropy) %>%
  gather(key = "Metric", value = "Value", 2:3) %>%
  mutate(Metric = ifelse(Metric == "Average_Type_Entropy", "Type Entropy", "Species Entropy")) %>%
  mutate(Region = ifelse(Site <= 915, "Structural", "Nonstructural")) %>%
  mutate(Metric_Region = paste0(Metric, ", ", Region))

write_csv(Type_EVC_1_Entropy_Comparison, "EVC_Entropy_Data.csv")

Type_EVC_1_Entropy_Comparison$Metric <- factor(Type_EVC_1_Entropy_Comparison$Metric, levels = c("Type Entropy", "Species Entropy"))

pdf("EVC_1_Alignment_Entropy_Comparison.pdf", height = 3.5, width = 8.5)
#png("EVC_1_Alignment_Entropy_Comparison.png", height = 1200, width = 2000, res = 300)

ggplot() +
  geom_rect(data = EVC_AA_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.6, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ma(data = EVC_Type_Shannon_Entropy, aes(x = Site, y = Average_Type_Entropy - SE_Type_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "lightgoldenrod", alpha = 0.5) +
  geom_ma(data = EVC_Type_Shannon_Entropy, aes(x = Site, y = Average_Type_Entropy + SE_Type_Entropy), ma_fun = SMA, n = 21, linetype = 1, color = "lightgoldenrod", alpha = 0.5) +
  geom_ma(data = EVC_1_Shannon_Entropy_Repeated, aes(x = Site, y = EVC_1_Entropy, group = Iteration), ma_fun = SMA, n = 21, linetype = 1, color = "goldenrod", alpha = 0.015) +
  geom_ma(data = Type_EVC_1_Entropy_Comparison, aes(x = Site, y = Value, color = Metric), ma_fun = SMA, n = 21, linetype = 1) +
  scale_color_manual(values = c("lightgoldenrod", "goldenrod")) +
  geom_vline(xintercept = 915, linetype = 2) +
  theme_prism() +
  xlab("Site (21aa window)") +
  ylab("Entropy") +
  theme(legend.position = c(1, 0.95), legend.justification = c("right", "top"), legend.margin = margin(0, 0, 0, 0), legend.text = element_text(size = 10)) +
  guides(fill = "none")

dev.off()

Type_EVC_1_Entropy_Comparison_Pvals = dunn_test(Value ~ Metric_Region, data = Type_EVC_1_Entropy_Comparison)

Type_EVC_1_Entropy_Comparison$Metric_Region = factor(Type_EVC_1_Entropy_Comparison$Metric_Region, levels = c("Type Entropy, Structural", "Type Entropy, Nonstructural", "Species Entropy, Structural", "Species Entropy, Nonstructural"))

pdf("Type_EVC_1_Entropy_Comparison_eCDF.pdf", height = 3.5, width = 5)

ggplot(Type_EVC_1_Entropy_Comparison, aes(x = Value, color = Metric_Region, linetype = Metric_Region)) +
  stat_ecdf(linewidth = 1) +
  scale_color_manual(values = c("lightgoldenrod", "lightgoldenrod", "goldenrod", "goldenrod")) +
  scale_linetype_manual(values = c(1, 2, 1, 2)) +
  xlim(1, 0) +
  xlab("Entropy") +
  ylab("Cumulative Proportion") +
  theme_prism() +
  theme(legend.position = c(.05, .95), legend.justification = c("left", "top"), legend.margin = margin(-3, -2, -2, -2), legend.text = element_text(size = 10)) +
  annotate("segment", x = c(0.40, 0.3985, 0.3985, 0.40), xend = c(0.40, 0.41, 0.41, 0.39), y = c(0.97, 0.97, 0.86, 0.915), yend = c(0.86, 0.97, 0.86, 0.915), colour = "black", size = 0.66) +
  annotate("text", x = 0.255, y = 0.92, label = "p = 6.110e-08", size = 3.5, fontface = "bold") +
  annotate("segment", x = c(0.35, 0.3485, 0.3485, 0.35), xend = c(0.35, 0.36, 0.36, 0.34), y = c(0.76, 0.76, 0.66, 0.705), yend = c(0.66, 0.76, 0.66, 0.705), colour = "black", size = 0.66) +
  annotate("text", x = 0.21, y = 0.71, label = "p = 7.402e-33", size = 3.5, fontface = "bold")

dev.off()
```

#Similarity calculations

Whole Genome
```{r}
EVC_Hamming_Distances_Matrix = DistanceMatrix(EVC_0.95_Codon_Alignment, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Hamming_Distances_DF = as.data.frame(EVC_Hamming_Distances_Matrix) %>%
  mutate(Seq_1 = rownames(EVC_Hamming_Distances_Matrix)) %>%
  gather(key = Seq_2, value = Hamming_Distance, 1:ncol(EVC_Hamming_Distances_Matrix)) %>%
  left_join(EVC_0.95_Types, join_by("Seq_1" == "name")) %>%
  mutate(Seq_1_Type = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  left_join(EVC_0.95_Types, join_by("Seq_2" == "name")) %>%
  mutate(Seq_2_Type = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  mutate(Comparison = ifelse(Seq_1_Type == Seq_2_Type, "Within type", "Between types"))

write_csv(EVC_Hamming_Distances_DF, "EVC_Hamming_Distances.csv")

EVC_Similarity_Summary = EVC_Hamming_Distances_DF %>%
  mutate(Similarity = 1 - Hamming_Distance) %>%
  summarise(Mean = mean(Similarity), SD = sd(Similarity), Min = min(Similarity), Max = max(Similarity))

ggplot(EVC_Hamming_Distances_DF, aes(x = Hamming_Distance, pattern = Comparison)) +
  geom_histogram_pattern(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7, pattern_color = "white", pattern_fill = "white", pattern_alpha = 1, pattern_angle = 45, pattern_spacing = 0.02, pattern_size = 0.1) +
  scale_pattern_manual(values = c(`Within type` = "stripe", `Between types` = "none")) +
  ylab("Count") +
  xlab("Distance") +
  theme_prism()
```

Sliding Window - EVC (nt)
```{r}
window_size = 200
overlap = 100

Average_Distance = NULL
i = 1

while (i < 6798) {
  end = ifelse((i + window_size) < 6798, i + window_size, 6798)
  
  EVC_Alignment_Segment = subseq(EVC_0.95_Codon_Alignment, i, end)
  
  EVC_Alignment_Segment_Distance = as.data.frame(DistanceMatrix(EVC_Alignment_Segment, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE))
  
  Seq_1 <- rownames(EVC_Alignment_Segment_Distance)
  
  EVC_Alignment_Segment_Distance = cbind(Seq_1, EVC_Alignment_Segment_Distance) %>%
    gather(key = Seq_2, value = Distance, 2:(ncol(EVC_Alignment_Segment_Distance) + 1)) %>%
    left_join(EVC_0.95_Types, join_by(Seq_1 == name)) %>%
    mutate(Type_1 = type) %>%
    select(-type, -`VP1 subgenogroup`) %>%
    left_join(EVC_0.95_Types, join_by(Seq_2 == name)) %>%
    mutate(Type_2 = type) %>%
    select(-type, -`VP1 subgenogroup`)
  
  EVC_Alignment_Segment_Distance_Within = EVC_Alignment_Segment_Distance %>%
    filter(Type_1 == Type_2) %>%
    filter(Seq_1 != Seq_2)
  
  EVC_Alignment_Segment_Distance_Between = EVC_Alignment_Segment_Distance %>%
    filter(Type_1 != Type_2)
  
  Average_Segment_Distance_Within = mean(EVC_Alignment_Segment_Distance_Within$Distance)
  SD_Segment_Distance_Within = sd(EVC_Alignment_Segment_Distance_Within$Distance)
  SE_Segment_Distance_Within = sd(EVC_Alignment_Segment_Distance_Within$Distance)/sqrt(length(EVC_0.95_Codon_Alignment))
  
  Average_Segment_Distance_Between = mean(EVC_Alignment_Segment_Distance_Between$Distance)
  SD_Segment_Distance_Between = sd(EVC_Alignment_Segment_Distance_Between$Distance)
  SE_Segment_Distance_Between = sd(EVC_Alignment_Segment_Distance_Between$Distance)/sqrt(length(EVC_0.95_Codon_Alignment))
  
  Segment_Entry = c(i, end, Average_Segment_Distance_Within, SD_Segment_Distance_Within, SE_Segment_Distance_Within, Average_Segment_Distance_Between, SD_Segment_Distance_Between, SE_Segment_Distance_Between)
  
  Average_Distance = rbind(Average_Distance, Segment_Entry)
  
  i = i + overlap
}

EVC_Average_Distance = as.data.frame(Average_Distance) %>%
  mutate(Scope = "EVC")
  
colnames(EVC_Average_Distance) <- c("Start", "End", "Within_Type_Distance", "Within_Type_SD", "Within_Type_SE", "Between_Type_Distance", "Between_Type_SD", "Between_Type_SE", "Scope")

pdf("EVC_Average_Distance.pdf", height = 3.5, width = 8.5)
#png("EVC_1_Alignment_Entropy_Comparison.png", height = 1200, width = 2000, res = 300)

ggplot() +
  geom_rect(data = EVC_Alignment_Features, aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.6, fill = Fill), alpha = 0.33) +
  scale_fill_manual(values = c("white", "lightgray")) +
  geom_ribbon(data = EVC_Average_Distance, aes(x = Start + 100, ymin = Within_Type_Distance - Within_Type_SD, ymax = Within_Type_Distance + Within_Type_SD), fill = "lightgoldenrod", alpha = 0.33) +
  geom_line(data = EVC_Average_Distance, aes(x = Start + 100, y = Within_Type_Distance), linetype = 1, color = "lightgoldenrod") +
  geom_ribbon(data = EVC_Average_Distance, aes(x = Start + 100, ymin = Between_Type_Distance - Between_Type_SD, ymax = Between_Type_Distance + Between_Type_SD), fill = "goldenrod", alpha = 0.33) +
  geom_line(data = EVC_Average_Distance, aes(x = Start + 100, y = Between_Type_Distance), linetype = 1, color = "goldenrod") +
  geom_vline(xintercept = 2875, linetype = 2) +
  theme_prism() +
  xlab("Site (nt)") +
  ylab("% Dissimilarity (nt)") +
  theme(legend.position = "none")

dev.off()

EVC_Average_Distance_Gathered = EVC_Average_Distance %>%
  select(Start, End, Within_Type_Distance, Between_Type_Distance) %>%
  mutate(`Within type` = Within_Type_Distance) %>%
  mutate(`Between types` = Between_Type_Distance) %>%
  select(-Within_Type_Distance, -Between_Type_Distance) %>%
  gather(key = Scope, value = Distance, 3:4)

EVC_Average_Distance_Gathered$Scope = factor(EVC_Average_Distance_Gathered$Scope, levels = c("Within type", "Between types"))

pdf("EVC_Average_Distance_LEGEND.pdf", height = 3.5, width = 7)

ggplot(EVC_Average_Distance_Gathered, aes(x = Start + 100, y = Distance, color = Scope)) +
  geom_line() +
  geom_vline(xintercept = 2875, linetype = 2) +
  theme_prism() +
  scale_color_manual(values = c("lightgoldenrod", "goldenrod")) +
  theme(legend.position = "right")

dev.off()
```

Structural vs. Nonstructural 
```{r}
EVC_Alignment_Structural = subseq(EVC_0.95_Codon_Alignment, 1, 2874)
EVC_Alignment_Structural_Distance = as.data.frame(DistanceMatrix(EVC_Alignment_Structural, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE))
EVC_Alignment_Structural_Distance_DF = cbind(Seq_1, EVC_Alignment_Structural_Distance) %>%
  gather(key = Seq_2, value = Distance, 2:(ncol(EVC_Alignment_Structural_Distance) + 1)) %>%
  mutate(Region = "Structural") %>%
  left_join(EVC_0.95_Types, join_by(Seq_1 == name)) %>%
  mutate(Type_1 = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  left_join(EVC_0.95_Types, join_by(Seq_2 == name)) %>%
  mutate(Type_2 = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  filter(Seq_1 != Seq_2) %>%
  mutate(Comparison = ifelse(Type_1 == Type_2, "Within type", "Between types"))

EVC_Alignment_Nonstructural = subseq(EVC_0.95_Codon_Alignment, 2875, 6798)
EVC_Alignment_Nonstructural_Distance = as.data.frame(DistanceMatrix(EVC_Alignment_Nonstructural, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE))
EVC_Alignment_Nonstructural_Distance_DF = cbind(Seq_1, EVC_Alignment_Nonstructural_Distance) %>%
  gather(key = Seq_2, value = Distance, 2:(ncol(EVC_Alignment_Nonstructural_Distance) + 1)) %>%
  mutate(Region = "Nonstructural") %>%
  left_join(EVC_0.95_Types, join_by(Seq_1 == name)) %>%
  mutate(Type_1 = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  left_join(EVC_0.95_Types, join_by(Seq_2 == name)) %>%
  mutate(Type_2 = type) %>%
  select(-type, -`VP1 subgenogroup`) %>%
  filter(Seq_1 != Seq_2) %>%
  mutate(Comparison = ifelse(Type_1 == Type_2, "Within type", "Between types"))

EVC_Alignment_Region_Distances = rbind(EVC_Alignment_Structural_Distance_DF, EVC_Alignment_Nonstructural_Distance_DF) %>%
  mutate(Metric = "EVC_nt", Species = "EVC") %>%
  mutate(Comparison_Region = paste0(Comparison, ", ", Region))

write_csv(EVC_Alignment_Region_Distances, "EVC_Region_Distances_Combined.csv")

EVC_Alignment_Region_Distances$Comparison = factor(EVC_Alignment_Region_Distances$Comparison, levels = c("Within type", "Between types"))

pdf("EVC_Alignment_Region_Distances.pdf", height = 3.5, width = 7)
#png("EVC_Alignment_Region_Distances.png", height = 1000, width = 2000, res = 300)

ggviolin(EVC_Alignment_Region_Distances, x = "Region", y = "Distance", fill = "Comparison", color = "black", add = "boxplot", add.params = list(fill = "white", width = 0.05)) +
  stat_compare_means(label.y = 0.45) +
  theme_prism() +
  ylab("% Dissimilarity (nt)") +
  scale_fill_manual(values = c("lightgoldenrod", "goldenrod")) +
  theme(legend.position = "none") +
  facet_wrap(~Comparison)

dev.off()

EVC_Alignment_Region_Distances$Comparison_Region = factor(EVC_Alignment_Region_Distances$Comparison_Region, level = c("Within type, Structural", "Within type, Nonstructural", "Between types, Structural", "Between types, Nonstructural"))

pdf("EVC_Dissimilarity_eCDF.pdf", height = 3.5, width = 5)

ggplot(EVC_Alignment_Region_Distances, aes(x = Distance, color = Comparison_Region, linetype = Comparison_Region)) +
  stat_ecdf(linewidth = 1) +
  scale_color_manual(values = c("lightgoldenrod", "lightgoldenrod", "goldenrod", "goldenrod")) +
  scale_linetype_manual(values = c(1, 2, 1, 2)) +
  xlim(1, 0) +
  xlab("% Dissimilarity (nt)") +
  ylab("Cumulative Proportion") +
  theme_prism() +
  theme(legend.position = c(.05, .95), legend.justification = c("left", "top"), legend.margin = margin(-3, -2, -2, -2), legend.text = element_text(size = 10))
  #annotate("segment", x = c(0.63, 0.6285, 0.6285, 0.63), xend = c(0.63, 0.64, 0.64, 0.62), y = c(0.97, 0.97, 0.90, 0.935), yend = c(0.90, 0.97, 0.90, 0.935), colour = "black", size = 0.66) +
  #annotate("text", x = 0.55, y = 0.94, label = "p < 2.2e-16", size = 3.5, fontface = "bold") +
  #annotate("segment", x = c(0.59, 0.5885, 0.5885, 0.59), xend = c(0.59, 0.60, 0.60, 0.58), y = c(0.83, 0.83, 0.76, 0.795), yend = c(0.76, 0.83, 0.76, 0.795), colour = "black", size = 0.66) +
  #annotate("text", x = 0.505, y = 0.80, label = "p < 2.2e-16", size = 3.5, fontface = "bold")

dev.off()
```

#Network construction
P1, P2, + P3 Clusters - BEAST Curated 95% Clustered 
```{r}
EVC_Alignment_P1 = subseq(EVC_0.95_Codon_Alignment, 1, 2874)
EVC_Alignment_P1_Similarity_Matrix = DistanceMatrix(EVC_Alignment_P1, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_P1_Similarity = as.data.frame(EVC_Alignment_P1_Similarity_Matrix)
writeXStringSet(EVC_Alignment_P1, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P1.fasta")

EVC_Alignment_P2 = subseq(EVC_0.95_Codon_Alignment, 2875, 4608)
EVC_Alignment_P2_Similarity_Matrix = DistanceMatrix(EVC_Alignment_P2, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_P2_Similarity = as.data.frame(EVC_Alignment_P2_Similarity_Matrix)
writeXStringSet(EVC_Alignment_P2, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P2.fasta")

EVC_Alignment_P3 = subseq(EVC_0.95_Codon_Alignment, 4609, 6798)
EVC_Alignment_P3_Similarity_Matrix = DistanceMatrix(EVC_Alignment_P3, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_P3_Similarity = as.data.frame(EVC_Alignment_P3_Similarity_Matrix)
writeXStringSet(EVC_Alignment_P3, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P3.fasta")
```

```{r}
EVC_Alignment_P1_Similarity_Dist = DistanceMatrix(EVC_Alignment_P1, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_P1_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_P1_Similarity_Dist)
colnames(EVC_Alignment_P1_Similarity_Dist_Summary) <- c("Distance")

pdf("EVC_Alignment_P1_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_P1_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2, alpha = 0.2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2) +
  xlim(-0.01, 0.43) +
  ylim(0, 3500) +
  ggtitle("P1") +
  ylab("Count") +
  xlab("Distance") +
  theme_prism()

#dev.off()

name <- rownames(EVC_Alignment_P1_Similarity_Matrix)

EVC_P1_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_P1_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_P1_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

```{r}
EVC_Alignment_P2_Similarity_Dist = DistanceMatrix(EVC_Alignment_P2, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_P2_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_P2_Similarity_Dist)
colnames(EVC_Alignment_P2_Similarity_Dist_Summary) <- c("Distance")

pdf("EVC_Alignment_P2_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_P2_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2, alpha = 0.2) +
  xlim(-0.01, 0.43) +
  ylim(0, 3500) +
  ggtitle("P2") +
  ylab("Count") + 
  xlab("Distance") +
  theme_prism()

dev.off()

name <- rownames(EVC_Alignment_P2_Similarity_Matrix)

EVC_P2_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_P2_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_P2_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

```{r}
EVC_Alignment_P3_Similarity_Dist = DistanceMatrix(EVC_Alignment_P3, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_P3_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_P3_Similarity_Dist)
colnames(EVC_Alignment_P3_Similarity_Dist_Summary) <- c("Distance")

pdf("EVC_Alignment_P3_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_P3_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2, alpha = 0.2) +
  xlim(-0.01, 0.43) +
  ylim(0, 3500) +
  ggtitle("P3") +
  ylab("Count") + 
  xlab("Distance") +
  theme_prism()

dev.off()

name <- rownames(EVC_Alignment_P3_Similarity_Matrix)

EVC_P3_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_P3_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_P3_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

Command line analyses:
1) Run cd-hit with c = 0.87 on All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P2.fasta [All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P2_0.87.fasta]
2) Run cd-hit with c = 0.87 on All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P3.fasta [All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P3_0.87.fasta]

Cluster Network - Type + P2/3 87% Clustering
```{r}
EVC_P2_Clusters = read_csv("Clustering_0.95/P1_P2_P3/P2_Clustering/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Trimmed_wDates_P2_0.87.fasta.bak.clstr", col_names = c("Col1", "Col2")) %>%
  separate(Col1, into = c("P2_Cluster", "Length")) %>%
  mutate(Accession = substr(Col2, 2, 11)) %>%
  mutate(Accession = ifelse((Accession == "V01149.1_1" | Accession == "D90457.1_1" | Accession == "K01392.1_1"), Accession, paste0(Accession, "_1"))) %>%
  select(P2_Cluster, Accession)

EVC_P3_Clusters = read_csv("Clustering_0.95/P1_P2_P3/P3_Clustering/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Trimmed_wDates_P3_0.87.fasta.bak.clstr", col_names = c("Col1", "Col2")) %>%
  separate(Col1, into = c("P3_Cluster", "Length")) %>%
  mutate(Accession = substr(Col2, 2, 11)) %>%
  mutate(Accession = ifelse((Accession == "V01149.1_1" | Accession == "D90457.1_1" | Accession == "K01392.1_1"), Accession, paste0(Accession, "_1"))) %>%
  select(P3_Cluster, Accession)
```

```{r}
EVC_WG_Groups = full_join(EVC_P2_Clusters, EVC_P3_Clusters, by = "Accession") %>%
  left_join(EVC_0.95_Types, join_by("Accession" == "name")) %>%
  select(Accession, type, P2_Cluster, P3_Cluster) %>%
  mutate(WG_Group = paste(type, P2_Cluster, P3_Cluster, sep = " | ")) %>%
  left_join(EVC_GeoLocations, by = "Accession") %>%
  separate(Geo_Location, ":", into = c("Geo_Location"))

EVC_WG_Groups_P1_P2 = EVC_WG_Groups %>%
  select(type, P2_Cluster) %>%
  mutate(Source = type) %>%
  mutate(Target = paste0("P2-", P2_Cluster)) %>%
  group_by(Source, Target) %>%
  summarise(n = n())
  
EVC_WG_Groups_P2_P3 = EVC_WG_Groups %>%
  select(P2_Cluster, P3_Cluster) %>%
  mutate(Source = paste0("P2-", P2_Cluster)) %>%
  mutate(Target = paste0("P3-", P3_Cluster)) %>%
  group_by(Source, Target) %>%
  summarise(n = n())
 
EVC_WG_Groups_Network = rbind(EVC_WG_Groups_P1_P2, EVC_WG_Groups_P2_P3) %>%
  select(Source, Target)
write_csv(EVC_WG_Groups_Network, "Clustering_0.95/P1_P2_P3/EVC_WG_Groups_Network.csv")

EVC_WG_Groups_P1_Node_Info = EVC_WG_Groups %>%
  group_by(type) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = type) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P1")

EVC_WG_Groups_P2_Node_Info = EVC_WG_Groups %>%
  group_by(P2_Cluster) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = paste0("P2-", P2_Cluster)) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P2")

EVC_WG_Groups_P3_Node_Info = EVC_WG_Groups %>%
  group_by(P3_Cluster) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = paste0("P3-", P3_Cluster)) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P3")

EVC_WG_Groups_Node_Info = rbind(EVC_WG_Groups_P1_Node_Info, EVC_WG_Groups_P2_Node_Info, EVC_WG_Groups_P3_Node_Info) %>%
  mutate(Species = "EVC") %>%
  mutate(Color_Indiciator = paste(Species, Region, sep = "-")) %>%
  mutate(nLocation = ifelse(nLocation == 0, 1, nLocation))
write_csv(EVC_WG_Groups_Node_Info, "Clustering_0.95/P1_P2_P3/EVC_WG_Groups_Node_Info.csv")

EVC_WG_Groups_Edge_Info = rbind(EVC_WG_Groups_P1_P2, EVC_WG_Groups_P2_P3) %>%
  ungroup() %>%
  mutate(Edge_Name = paste0(Source, " (interacts with) ", Target)) %>%
  select(Edge_Name, n)
write_csv(EVC_WG_Groups_Edge_Info, "Clustering_0.95/P1_P2_P3/EVC_WG_Groups_Edge_Info.csv")
```

```{r}
Type_Code <- c(1:20)

EVC_Types_Code = EVC_WG_Groups %>%
  distinct(type) %>%
  arrange(type) %>%
  cbind(Type_Code)

EVC_WG_Groups_Gathered = EVC_WG_Groups %>%
  left_join(EVC_Types_Code, by = "type") %>%
  select(Accession, type, Type_Code, P2_Cluster, P3_Cluster) %>%
  gather(key = "Region", value = "Cluster", 3:5)

EVC_WG_Groups_Gathered$Region <- factor(EVC_WG_Groups_Gathered$Region, levels = c("Type_Code", "P2_Cluster", "P3_Cluster"))

ggplot(EVC_WG_Groups_Gathered, aes(x = Region, y = Cluster, color = type, group = Accession)) +
  geom_point() +
  geom_line() +
  theme_prism() +
  theme(legend.position = "none")
```

Visualize networks in Cytoscape and extract network statistics

Network Statistics
```{r}
EVC_Network_Stats = read_csv("Clustering_0.95/P1_P2_P3/EVC_Network_Statistics.csv")  %>%
  mutate(Dataset = "Without\nVaccine-derived\nSequences")

mean(EVC_Network_Stats$NeighborhoodConnectivity)
```

#Network construction - w/ vaccine sequences
P1, P2, + P3 Clusters - BEAST Curated 95% Clustered 
```{r}
EVC_Alignment_wVaccine_P1 = subseq(EVC_0.95_Codon_Alignment_wVaccine, 1, 2874)
EVC_Alignment_wVaccine_P1_Similarity_Matrix = DistanceMatrix(EVC_Alignment_wVaccine_P1, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_wVaccine_P1_Similarity = as.data.frame(EVC_Alignment_wVaccine_P1_Similarity_Matrix)
writeXStringSet(EVC_Alignment_wVaccine_P1, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_wVaccine_P1.fasta")

EVC_Alignment_wVaccine_P2 = subseq(EVC_0.95_Codon_Alignment_wVaccine, 2875, 4608)
EVC_Alignment_wVaccine_P2_Similarity_Matrix = DistanceMatrix(EVC_Alignment_wVaccine_P2, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_wVaccine_P2_Similarity = as.data.frame(EVC_Alignment_wVaccine_P2_Similarity_Matrix)
writeXStringSet(EVC_Alignment_wVaccine_P2, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_wVaccine_P2.fasta")

EVC_Alignment_wVaccine_P3 = subseq(EVC_0.95_Codon_Alignment_wVaccine, 4609, 6798)
EVC_Alignment_wVaccine_P3_Similarity_Matrix = DistanceMatrix(EVC_Alignment_wVaccine_P3, type = "matrix", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)
EVC_Alignment_wVaccine_P3_Similarity = as.data.frame(EVC_Alignment_wVaccine_P3_Similarity_Matrix)
writeXStringSet(EVC_Alignment_wVaccine_P3, "Clustering_0.95/P1_P2_P3/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_wVaccine_P3.fasta")
```

```{r}
EVC_Alignment_wVaccine_P1_Similarity_Dist = DistanceMatrix(EVC_Alignment_wVaccine_P1, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_wVaccine_P1_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_wVaccine_P1_Similarity_Dist)
colnames(EVC_Alignment_wVaccine_P1_Similarity_Dist_Summary) <- c("Distance")

#pdf("EVC_Alignment_wVaccine_P1_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_wVaccine_P1_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2) +
  xlim(-0.01, 0.43) +
  #ylim(0, 3500) +
  ggtitle("P1") +
  ylab("Count") +
  xlab("Distance") +
  theme_prism()

#dev.off()

name <- rownames(EVC_Alignment_wVaccine_P1_Similarity_Matrix)

EVC_wVaccine_P1_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_wVaccine_P1_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_wVaccine_P1_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

```{r}
EVC_Alignment_wVaccine_P2_Similarity_Dist = DistanceMatrix(EVC_Alignment_wVaccine_P2, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_wVaccine_P2_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_wVaccine_P2_Similarity_Dist)
colnames(EVC_Alignment_wVaccine_P2_Similarity_Dist_Summary) <- c("Distance")

#pdf("EVC_Alignment_wVaccine_P2_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_wVaccine_P2_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2) +
  xlim(-0.01, 0.43) +
  #ylim(0, 3500) +
  ggtitle("P2") +
  ylab("Count") + 
  xlab("Distance") +
  theme_prism()

#dev.off()

name <- rownames(EVC_Alignment_wVaccine_P2_Similarity_Matrix)

EVC_wVaccine_P2_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_wVaccine_P2_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_wVaccine_P2_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

```{r}
EVC_Alignment_wVaccine_P3_Similarity_Dist = DistanceMatrix(EVC_Alignment_wVaccine_P3, type = "dist", includeTerminalGaps = FALSE, correction = "none", verbose = TRUE)

EVC_Alignment_wVaccine_P3_Similarity_Dist_Summary = as.data.frame(EVC_Alignment_wVaccine_P3_Similarity_Dist)
colnames(EVC_Alignment_wVaccine_P3_Similarity_Dist_Summary) <- c("Distance")

#pdf("EVC_Alignment_wVaccine_P3_Similarity.pdf", height = 4, width = 6)

ggplot(EVC_Alignment_wVaccine_P3_Similarity_Dist_Summary, aes(x = Distance)) +
  geom_histogram(binwidth = 0.01, fill = "goldenrod", color = "black", alpha = 0.7) +
  geom_vline(xintercept = 0.13, color = "black", linetype = 2) +
  geom_vline(xintercept = 0.25, color = "black", linetype = 2) +
  xlim(-0.01, 0.43) +
  #ylim(0, 3500) +
  ggtitle("P3") +
  ylab("Count") + 
  xlab("Distance") +
  theme_prism()

#dev.off()

name <- rownames(EVC_Alignment_wVaccine_P3_Similarity_Matrix)

EVC_wVaccine_P3_MSD = cbind(name, as.data.frame(cmdscale(EVC_Alignment_wVaccine_P3_Similarity_Matrix))) %>%
  left_join(EVC_0.95_Types, by = "name")

ggplot(EVC_wVaccine_P3_MSD, aes(x = V1, y = V2, color = type)) +
  geom_point() +
  theme_prism()
```

Command line analyses:
1) Run cd-hit with c = 0.87 on All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P2.fasta [All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_wVaccine_P2_0.87.fasta]
2) Run cd-hit with c = 0.87 on All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_P3.fasta [All_Enterovirus_C_ORF_Curated_Sequences_0.95_Codon_Aligned_Trimmed_wDates_wVaccine_P3_0.87.fasta]

Cluster Network - Type + P2/3 87% Clustering
```{r}
EVC_wVaccine_P2_Clusters = read_csv("Clustering_0.95/P1_P2_P3/P2_Clustering/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Trimmed_wDates_wVaccine_P2_0.87.fasta.bak.clstr", col_names = c("Col1", "Col2")) %>%
  separate(Col1, into = c("P2_Cluster", "Length")) %>%
  mutate(Accession = substr(Col2, 2, 11)) %>%
  mutate(Accession = ifelse((Accession == "V01149.1_1" | Accession == "D90457.1_1" | Accession == "K01392.1_1"), Accession, paste0(Accession, "_1"))) %>%
  select(P2_Cluster, Accession)

EVC_wVaccine_P3_Clusters = read_csv("Clustering_0.95/P1_P2_P3/P3_Clustering/All_Enterovirus_C_ORF_Curated_Sequences_0.95_Trimmed_wDates_wVaccine_P3_0.87.fasta.bak.clstr", col_names = c("Col1", "Col2")) %>%
  separate(Col1, into = c("P3_Cluster", "Length")) %>%
  mutate(Accession = substr(Col2, 2, 11)) %>%
  mutate(Accession = ifelse((Accession == "V01149.1_1" | Accession == "D90457.1_1" | Accession == "K01392.1_1"), Accession, paste0(Accession, "_1"))) %>%
  select(P3_Cluster, Accession)
```

```{r}
EVC_wVaccine_WG_Groups = full_join(EVC_wVaccine_P2_Clusters, EVC_wVaccine_P3_Clusters, by = "Accession") %>%
  left_join(EVC_0.95_Types, join_by("Accession" == "name")) %>%
  select(Accession, type, P2_Cluster, P3_Cluster) %>%
  mutate(WG_Group = paste(type, P2_Cluster, P3_Cluster, sep = " | ")) %>%
  left_join(EVC_GeoLocations, by = "Accession") %>%
  separate(Geo_Location, ":", into = c("Geo_Location"))

EVC_wVaccine_WG_Groups_P1_P2 = EVC_wVaccine_WG_Groups %>%
  select(type, P2_Cluster) %>%
  mutate(Source = type) %>%
  mutate(Target = paste0("P2-", P2_Cluster)) %>%
  group_by(Source, Target) %>%
  summarise(n = n())
  
EVC_wVaccine_WG_Groups_P2_P3 = EVC_wVaccine_WG_Groups %>%
  select(P2_Cluster, P3_Cluster) %>%
  mutate(Source = paste0("P2-", P2_Cluster)) %>%
  mutate(Target = paste0("P3-", P3_Cluster)) %>%
  group_by(Source, Target) %>%
  summarise(n = n())
 
EVC_wVaccine_WG_Groups_Network = rbind(EVC_wVaccine_WG_Groups_P1_P2, EVC_wVaccine_WG_Groups_P2_P3) %>%
  select(Source, Target)
write_csv(EVC_wVaccine_WG_Groups_Network, "Clustering_0.95/P1_P2_P3/EVC_wVaccine_WG_Groups_Network.csv")

EVC_wVaccine_WG_Groups_P1_Node_Info = EVC_wVaccine_WG_Groups %>%
  group_by(type) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = type) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P1")

EVC_wVaccine_WG_Groups_P2_Node_Info = EVC_wVaccine_WG_Groups %>%
  group_by(P2_Cluster) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = paste0("P2-", P2_Cluster)) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P2")

EVC_wVaccine_WG_Groups_P3_Node_Info = EVC_wVaccine_WG_Groups %>%
  group_by(P3_Cluster) %>%
  summarise(Count = n(), nLocation = n_distinct(Geo_Location, na.rm = TRUE)) %>%
  mutate(Node = paste0("P3-", P3_Cluster)) %>%
  select(Node, Count, nLocation) %>%
  mutate(Region = "P3")

EVC_wVaccine_WG_Groups_Node_Info = rbind(EVC_wVaccine_WG_Groups_P1_Node_Info, EVC_wVaccine_WG_Groups_P2_Node_Info, EVC_wVaccine_WG_Groups_P3_Node_Info) %>%
  mutate(Species = "EVC") %>%
  mutate(Color_Indiciator = paste(Species, Region, sep = "-")) %>%
  mutate(nLocation = ifelse(nLocation == 0, 1, nLocation))
write_csv(EVC_wVaccine_WG_Groups_Node_Info, "Clustering_0.95/P1_P2_P3/EVC_wVaccine_WG_Groups_Node_Info.csv")

EVC_wVaccine_WG_Groups_Edge_Info = rbind(EVC_wVaccine_WG_Groups_P1_P2, EVC_wVaccine_WG_Groups_P2_P3) %>%
  ungroup() %>%
  mutate(Edge_Name = paste0(Source, " (interacts with) ", Target)) %>%
  select(Edge_Name, n)
write_csv(EVC_wVaccine_WG_Groups_Edge_Info, "Clustering_0.95/P1_P2_P3/EVC_wVaccine_WG_Groups_Edge_Info.csv")
```

```{r}
Type_Code <- c(1:20)

EVC_Types_Code = EVC_WG_Groups %>%
  distinct(type) %>%
  arrange(type) %>%
  cbind(Type_Code)

EVC_wVaccine_WG_Groups_Gathered = EVC_wVaccine_WG_Groups %>%
  left_join(EVC_Types_Code, by = "type") %>%
  select(Accession, type, Type_Code, P2_Cluster, P3_Cluster) %>%
  gather(key = "Region", value = "Cluster", 3:5)

EVC_wVaccine_WG_Groups_Gathered$Region <- factor(EVC_wVaccine_WG_Groups_Gathered$Region, levels = c("Type_Code", "P2_Cluster", "P3_Cluster"))

ggplot(EVC_wVaccine_WG_Groups_Gathered, aes(x = Region, y = Cluster, color = type, group = Accession)) +
  geom_point() +
  geom_line() +
  theme_prism() +
  theme(legend.position = "none")
```

Visualize networks in Cytoscape and extract network statistics

Network Statistics
```{r}
EVC_wVaccine_Network_Stats = read_csv("Clustering_0.95/P1_P2_P3/EVC_wVaccine_Network_Statistics.csv") %>%
  mutate(Dataset = "With\nVaccine-derived\nSequences")

mean(EVC_wVaccine_Network_Stats$NeighborhoodConnectivity)
median(EVC_wVaccine_Network_Stats$NeighborhoodConnectivity)
```

Network Statistics Comparison: with + without vaccine sequences
```{r}
EVC_Network_Stats_Comparison = rbind(EVC_Network_Stats, EVC_wVaccine_Network_Stats)

EVC_Network_Stats_Comparison$Dataset <- factor(EVC_Network_Stats_Comparison$Dataset, levels = c("Without\nVaccine-derived\nSequences", "With\nVaccine-derived\nSequences"))

pdf("EVC_Network_Stats_Comparison.pdf", height = 7, width = 4)

ggplot(EVC_Network_Stats_Comparison, aes(x = Dataset, y = NeighborhoodConnectivity)) +
  stat_compare_means(label.x = 0.8, label.y = 22) +
  geom_boxplot_pattern(fill = "goldenrod", aes(pattern = Dataset), pattern_color = "white", pattern_fill = "white", pattern_alpha = 1, pattern_angle = 45, pattern_spacing = 0.04, pattern_size = 0.1) +
  scale_pattern_manual(values = c(`With\nVaccine-derived\nSequences` = "stripe", `Without\nVaccine-derived\nSequences` = "none")) +
  theme_prism() +
  xlab("") +
  ylab("Neighborhood Connectivity") +
  theme(legend.position = "none", axis.text.x = element_text(size = 12))

dev.off()
```

Network Statistics Comparison: with + without vaccine sequences by 3Dpol group
```{r}
EVC_PolGroup_I_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_PolGroup_I_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "I") %>%
  mutate(Dataset = "No")
EVC_PolGroup_II_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_PolGroup_II_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "II") %>%
  mutate(Dataset = "No")
EVC_PolGroup_III_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_PolGroup_III_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "III") %>%
  mutate(Dataset = "No")
EVC_PolGroup_IV_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_PolGroup_IV_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "IV") %>%
  mutate(Dataset = "No")

EVC_wVaccine_PolGroup_I_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_wVaccine_PolGroup_I_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "I") %>%
  mutate(Dataset = "Yes")
EVC_wVaccine_PolGroup_II_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_wVaccine_PolGroup_II_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "II") %>%
  mutate(Dataset = "Yes")
EVC_wVaccine_PolGroup_III_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_wVaccine_PolGroup_III_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "III") %>%
  mutate(Dataset = "Yes")
EVC_wVaccine_PolGroup_IV_Network_Statistics = read_csv("Clustering_0.95/P1_P2_P3/EVC_wVaccine_PolGroup_IV_Network_Statistics.csv") %>%
  select(Species, name, NeighborhoodConnectivity) %>%
  mutate(PolGroup = "IV") %>%
  mutate(Dataset = "Yes")

EVC_PolGroup_Network_Stats_Comparison = rbind(EVC_PolGroup_I_Network_Statistics, EVC_PolGroup_II_Network_Statistics, EVC_PolGroup_III_Network_Statistics, EVC_PolGroup_IV_Network_Statistics, EVC_wVaccine_PolGroup_I_Network_Statistics, EVC_wVaccine_PolGroup_II_Network_Statistics, EVC_wVaccine_PolGroup_III_Network_Statistics, EVC_wVaccine_PolGroup_IV_Network_Statistics)
```

```{r}
EVC_PolGroup_Network_Stats_Comparison$Dataset = factor(EVC_PolGroup_Network_Stats_Comparison$Dataset, levels = c("No", "Yes"))

pdf("EVC_PolGroup_Network_Stats_Comparison.pdf", height = 4, width = 10)

ggplot(EVC_PolGroup_Network_Stats_Comparison, aes(x = Dataset, y = NeighborhoodConnectivity, fill = PolGroup)) +
  geom_boxplot_pattern(aes(pattern = Dataset), color = "black", pattern_color = "white", pattern_fill = "white", pattern_alpha = 1, pattern_angle = 45, pattern_spacing = 0.04, pattern_size = 0.1) +
  scale_pattern_manual(values = c(`Yes` = "stripe", `No` = "none")) +
  scale_fill_manual(values = c("#f7020295", "#02f70a95", "#f7df0295", "#0213f795")) +
  theme_prism() +
  xlab("Inclusion of Vaccine-derived Sequences") +
  ylab("Neighborhood Connectivity") +
  theme(legend.position = "none", axis.text.x = element_text(size = 12), axis.title.x = ggtext::element_markdown()) +
  facet_wrap(~PolGroup, nrow = 1)

dev.off()
```

#SSE + TreeOrderScan

Command line analyses:
1) Reduce data to a managable amount for TreeOrderScan by subclustering at 98% on the amino acid sequences in All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_AA.fasta using cd-hit (c = 0.98) [All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_AA_0.98.fasta]

Add information to sequence names for parsing in SSE
```{r}
EVC_0.95_AA_0.98_Clusters = readAAStringSet("Clustering_0.95/Clustering_0.98/All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_woVaccine_AA_0.98.fasta")

Accession <- substr(names(EVC_0.95_AA_0.98_Clusters), 1, nchar(names(EVC_0.95_AA_0.98_Clusters)) - 2)

EVC_0.95_AA_0.98_ORFs = EVC_ORFs_Complete_Sequences[names(EVC_ORFs_Complete_Sequences) %in% Accession]

EVC_0.95_AA_0.98_names = as.data.frame(Accession) %>%
  left_join(EVC_WG_Groups, join_by("Accession")) %>%
  select(Accession, WG_Group) %>%
  mutate(name = paste(Accession, WG_Group, sep = " | "))

for (i in 1:length(EVC_0.95_AA_0.98_ORFs)) {
  names(EVC_0.95_AA_0.98_ORFs)[names(EVC_0.95_AA_0.98_ORFs) == EVC_0.95_AA_0.98_names$Accession[i]] <- EVC_0.95_AA_0.98_names$name[i]
}

writeXStringSet(EVC_0.95_AA_0.98_ORFs, "Clustering_0.95/Clustering_0.98/All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_woVaccine_AA_0.98_SSEnames.fasta")
```

Randomly select 50 sequences from the curated 95% clustered dataset for SSE
```{r}
set.seed(924)

EVC_SSE_Sample50 = sample(names(EVC_0.95_Codon_Alignment), 50)

EVC_0.95_ORFs_Sample50 = EVC_ORFs_Complete_Sequences[names(EVC_ORFs_Complete_Sequences) %in% EVC_SSE_Sample50]

EVC_SSE_Sample50_names = as.data.frame(EVC_SSE_Sample50) %>%
  dplyr::rename(Accession = EVC_SSE_Sample50) %>%
  left_join(EVC_WG_Groups, join_by("Accession")) %>%
  select(Accession, WG_Group) %>%
  mutate(name = paste(Accession, WG_Group, sep = " | "))

for (i in 1:length(EVC_0.95_ORFs_Sample50)) {
  names(EVC_0.95_ORFs_Sample50)[names(EVC_0.95_ORFs_Sample50) == EVC_SSE_Sample50_names$Accession[i]] <- EVC_SSE_Sample50_names$name[i]
}

writeXStringSet(EVC_0.95_ORFs_Sample50, "Clustering_0.95/SSE_Sample_50/All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_Sample50_SSEnames.fasta")
```

Run TreeOrderScan in SSE with the following parameters:
1) Number of bootstraps: 100
2) Bootstrap threshold: 70
3) Length of fragment: 200
4) Increment between fragments: 100
5) Group size reporting threshold: 1

Phylogenetic compatibility matrix
```{r}
EVC_TOS_ViolationList =  read_tsv("Clustering_0.95/Clustering_0.98/Enterovirus_C.Violation_List.DAT") %>%
  filter(Pos_1 != 6612)

pal <- wes_palette("Zissou1", 100, type = "continuous")

pdf("EVC_TOS_Compatibility_Matrix_AG.pdf", width = 4, height = 4)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `AG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 4)) +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_Compatibility_Matrix_AG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `AG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 4), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()

pdf("EVC_TOS_Compatibility_Matrix_BG.pdf", width = 4, height = 4)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `BG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 2), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_Compatibility_Matrix_BG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `BG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 2), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()

pdf("EVC_TOS_Compatibility_Matrix_WG.pdf", width = 4, height = 4)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `WG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 0.75), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_Compatibility_Matrix_WG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `WG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 0.75), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()
```

```{r}
EVC_TOS_ViolationFreq =  read_tsv("Clustering_0.95/Clustering_0.98/Enterovirus_C_Tagged.Violation_Freq.DAT")
```

Phylogenetic compatibility matrix - 50 samples
```{r}
EVC_50_TOS_ViolationList =  read_tsv("Clustering_0.95/Enterovirus_C.Violation_List.DAT") %>%
  filter(Pos_1 != 6654)

pal <- wes_palette("Zissou1", 100, type = "continuous")

pdf("EVC_50_TOS_Compatibility_Matrix_AG.pdf", width = 4, height = 4)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `AG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 4)) +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_50_Compatibility_Matrix_AG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `AG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 4), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()

pdf("EVC_50_TOS_Compatibility_Matrix_BG.pdf", width = 4, height = 4)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `BG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 1.75), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_Compatibility_Matrix_BG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `BG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 1.75), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()

pdf("EVC_50_TOS_Compatibility_Matrix_WG.pdf", width = 4, height = 4)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `WG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 0.5), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  ggtitle("Enterovirus C") +
  theme_prism() +
  theme(legend.position = "none")

dev.off()

pdf("TOS_Compatibility_Matrix_WG_LEGEND.pdf", width = 4, height = 5)

ggplot(EVC_50_TOS_ViolationList, aes(x = Pos_1, y = Pos_2, fill = `WG/clade`)) +
  geom_tile() +
  scale_fill_gradientn(colors = pal, limits = c(0, 0.5), name = "Violations \nper Clade\n") +
  xlab("") +
  ylab("") +
  theme_prism() +
  theme(legend.title = element_text(), legend.title.align=0.5)

dev.off()
```

```{r}
EVC_50_TOS_ViolationFreq =  read_tsv("Clustering_0.95/Enterovirus_C.Violation_Freq.DAT")
```

#RDP4 Results to prep for BEAST

Run RDP4 on All_Enterovirus_C_ORF_Curated_Sequences_Trimmed_CompleteSeqs_0.95_AA_0.98_Aligned_SSEnames.fasta, then download the alignment without recombinant regions [All_Enterovirus_C_0.95_AA_0.98_RDP4_NoRecombinantRegions.fas]

Extract 3D sequence
```{r}
EVC_0.95_AA_0.98_NoRecombinantRegions_Outgroup = readDNAStringSet("Clustering_0.95/Clustering_0.98/All_Enterovirus_C_0.95_AA_0.98_Outgroup_RDP4_NoRecombinantRegions.fas")

EVC_0.95_AA_0.98_NoRecombinantRegions_Outgroup_3D = subseq(EVC_0.95_AA_0.98_NoRecombinantRegions_Outgroup, 5371, 6759)
EVC_0.95_AA_0.98_NoRecombinantRegions_3D = EVC_0.95_AA_0.98_NoRecombinantRegions_Outgroup_3D[names(EVC_0.95_AA_0.98_NoRecombinantRegions_Outgroup_3D) != "FJ445111.1_1 | Rhinovirus A1"]
writeXStringSet(EVC_0.95_AA_0.98_NoRecombinantRegions_3D, "Clustering_0.95/Clustering_0.98/3D_tMRCA/All_Enterovirus_C_0.95_AA_0.98_RDP4_NoRecombinantRegions_3D.fasta")

Accession = names(EVC_0.95_AA_0.98_NoRecombinantRegions_3D)
Sequences = as.character(EVC_0.95_AA_0.98_NoRecombinantRegions_3D)

EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF = as.data.frame(cbind(Accession, Sequences)) %>%
  mutate(Gap_Count = str_count(Sequences, "-")) %>%
  mutate(Gap_Frequency = Gap_Count/width(EVC_0.95_AA_0.98_NoRecombinantRegions_3D)[1])

EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced = EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF %>%
  filter(Gap_Frequency < 0.1) %>%
  mutate(Accession_1 = Accession) %>%
  separate(Accession_1, c("Name"), " | ") %>%
  select(Accession, Name, Sequences) 

EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced_Dates = EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced %>% 
  select(Name) %>%
  left_join(EVC_Dates, join_by("Name" == "Accession")) %>%
  select(Name, BEAST_Date)

write_tsv(EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced_Dates, "Clustering_0.95/Clustering_0.98/3D_tMRCA/Curated_EVC_3D_Sequence_Info_Complete_Dates.tsv")

EVC_0.95_AA_0.98_NoRecombinantRegions_3D_Reduced = DNAStringSet(EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced$Sequences)
names(EVC_0.95_AA_0.98_NoRecombinantRegions_3D_Reduced) <- EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced$Name

writeXStringSet(EVC_0.95_AA_0.98_NoRecombinantRegions_3D_Reduced, "Clustering_0.95/Clustering_0.98/3D_tMRCA/All_Enterovirus_C_0.95_AA_0.98_RDP4_NoRecombinantRegions_3D_Reduced.fasta")
```

Extract 3D-linked VP1 sequences
```{r}
Accessions <- EVC_0.95_AA_0.98_NoRecombinantRegions_3D_DF_Reduced$Name

EVC_0.95_AA_0.98_ORFs_Codon_Aligned = EVC_0.95_Codon_Alignment[names(EVC_0.95_Codon_Alignment) %in% Accessions]
EVC_0.95_AA_0.98_ORFs_Codon_Aligned_VP1 = subseq(EVC_0.95_AA_0.98_ORFs_Codon_Aligned, 1834, 2874)

writeXStringSet(EVC_0.95_AA_0.98_ORFs_Codon_Aligned_VP1, "Clustering_0.95/Clustering_0.98/VP1_tMRCA/All_Enterovirus_C_0.95_AA_0.98_RDP4_NoRecombinantRegions_Reduced_VP1.fasta")
```

#BEAST (tMRCA)

Extract VP1 sequences for BEAST
```{r}
EVC_0.95_Codon_Alignment_VP1 = subseq(EVC_0.95_Codon_Alignment, 1834, 2874)
writeXStringSet(EVC_0.95_Codon_Alignment_VP1, "Clustering_0.95/VP1_tMRCA/All_Enterovirus_C_ORF_Curated_Sequences_Codon_Aligned_Trimmed_CompleteSeqs_0.95_VP1.fasta")
```

Tree Statistics
```{r}
EVC_VP1_Tree = read.nexus("Clustering_0.95/Clustering_0.98/VP1_3D_BEAST/BEAST_Partitioned_RootHeightLN2.0_Runs_Combined_VP1.trees")
EVC_3D_Tree = read.nexus("Clustering_0.95/Clustering_0.98/VP1_3D_BEAST/BEAST_Partitioned_RootHeightLN2.0_Runs_Combined_3D.trees")

EVC_TreeStats = NULL

for(i in 1:length(EVC_VP1_Tree)) {
  TEMP_VP1_Treeness = treeness(EVC_VP1_Tree[[i]])
  TEMP_3D_Treeness = treeness(EVC_3D_Tree[[i]])
  
  TEMP_RF_Distance = RF.dist(EVC_VP1_Tree[[i]], EVC_3D_Tree[[i]], check.labels = TRUE, rooted = TRUE, normalize = TRUE)
  
  TEMP_TreeStats = as.data.frame(cbind(i, TEMP_VP1_Treeness, TEMP_3D_Treeness, TEMP_RF_Distance))
  
  EVC_TreeStats = rbind(EVC_TreeStats, TEMP_TreeStats)
}

colnames(EVC_TreeStats) <- c("Run", "VP1_Treeness", "3D_Treeness", "RF_Distance")

write_csv(EVC_TreeStats, "EVC_Tree_Statistics.csv")
```

Tanglegram
```{r}
EVC_VP1_MCC_Tree = read.nexus("Clustering_0.95/Clustering_0.98/VP1_3D_BEAST/BEAST_Partitioned_RootHeightLN2.0_Runs_Combined_VP1.mcc.trees")
EVC_3D_MCC_Tree = read.nexus("Clustering_0.95/Clustering_0.98/VP1_3D_BEAST/BEAST_Partitioned_RootHeightLN2.0_Runs_Combined_3D.mcc.trees")

EVC_VP1_Tree_Dend <- ladder(as.dendrogram(EVC_VP1_MCC_Tree))
EVC_3D_Tree_Dend <- ladder(as.dendrogram(EVC_3D_MCC_Tree))

set.seed(3954)
EVC_Palette <- createPalette(20, c("#DAA520"), M=1000)
swatch(EVC_Palette)
EVC_Type_Colors = cbind(EVC_0.95_Types_Summary, EVC_Palette) %>%
  select(type, EVC_Palette)

name <- labels(EVC_VP1_Tree_Dend)
EVC_0.95_AA_0.98_Type_Colors = as.data.frame(name) %>%
  left_join(EVC_0.95_Types, by = "name") %>%
  left_join(EVC_Type_Colors, by = "type") %>%
  select(name, EVC_Palette)

labels_colors(EVC_VP1_Tree_Dend) <- EVC_0.95_AA_0.98_Type_Colors$EVC_Palette

name <- labels(EVC_3D_Tree_Dend)
EVC_0.95_AA_0.98_Type_Colors = as.data.frame(name) %>%
  left_join(EVC_0.95_Types, by = "name") %>%
  left_join(EVC_Type_Colors, by = "type") %>%
  select(name, EVC_Palette)

labels_colors(EVC_3D_Tree_Dend) <- EVC_0.95_AA_0.98_Type_Colors$EVC_Palette

EVC_tanglegram = untangle(EVC_VP1_Tree_Dend, EVC_3D_Tree_Dend, method = "stepBothSides") %>%
  tanglegram(highlight_distinct_edges = TRUE, highlight_branches_lwd = FALSE, margin_inner = 7, lwd = 2)

name <- labels(EVC_tanglegram$dend1)
EVC_0.95_AA_0.98_Type_Colors_New_Dend1 = as.data.frame(name) %>%
  left_join(EVC_0.95_Types, by = "name") %>%
  left_join(EVC_Type_Colors, by = "type") %>%
  select(name, EVC_Palette)

name <- labels(EVC_tanglegram$dend2)
EVC_0.95_AA_0.98_Type_Colors_New_Dend2 = as.data.frame(name) %>%
  left_join(EVC_0.95_Types, by = "name") %>%
  left_join(EVC_Type_Colors, by = "type") %>%
  select(name, EVC_Palette)

EVC_tanglegram_colored = untangle(EVC_VP1_Tree_Dend, EVC_3D_Tree_Dend, method = "stepBothSides") %>%
  tanglegram(highlight_distinct_edges = TRUE, highlight_branches_lwd = FALSE, margin_inner = 7, lwd = 2, color_lines = EVC_0.95_AA_0.98_Type_Colors_New_Dend1$EVC_Palette, columns_width = c(8.92, 4, 7.15))

EVC_tanglegram_colored$dend1 = EVC_tanglegram_colored$dend1 %>%
  set("leaves_pch", 21) %>%
  set("leaves_cex", 2) %>%
  set("leaves_col", "#000000") %>%
  set("leaves_bg", EVC_0.95_AA_0.98_Type_Colors_New_Dend1$EVC_Palette) %>%
  set("labels_col", "#FFFFFF")

EVC_tanglegram_colored$dend2 = EVC_tanglegram_colored$dend2 %>%
  set("leaves_pch", 21) %>%
  set("leaves_cex", 2) %>%
  set("leaves_col", "#000000") %>%
  set("leaves_bg", EVC_0.95_AA_0.98_Type_Colors_New_Dend2$EVC_Palette) %>%
  set("labels_col", "#FFFFFF")

pdf("EVC_Tanglegram.pdf", width = 9.8, height = 10.58)

plot(EVC_tanglegram_colored, highlight_distinct_edges = TRUE, edge.lwd = 2, margin_inner = 0, lwd = 2, color_lines = EVC_0.95_AA_0.98_Type_Colors_New_Dend1$EVC_Palette, columns_width = c(8.92, 4, 7.15), xaxs = "r", left_dendo_mar = c(2.5, 0.5, 3, 0), right_dendo_mar = c(2.5, 0.2, 3, 0.5))

dev.off()
```

LTT + GTT
```{r}
EVC_VP1_Tree_Ultra = force.ultrametric(EVC_VP1_MCC_Tree, method = "nnls")
EVC_3D_Tree_Ultra = force.ultrametric(EVC_3D_MCC_Tree, method = "nnls")

ltt(EVC_VP1_Tree_Ultra, log = TRUE)
lines(c(0, max(nodeHeights(EVC_VP1_Tree_Ultra))), c(log(2), log(length(EVC_VP1_Tree_Ultra$tip.label))),lty = "dashed", lwd = 2, col = "red")

ltt(EVC_VP1_MCC_Tree)

EVC_VP1_gtt = gtt(EVC_VP1_Tree_Ultra)
plot(EVC_VP1_gtt)

ltt(EVC_3D_Tree_Ultra, log = TRUE)
lines(c(0, max(nodeHeights(EVC_3D_Tree_Ultra))), c(log(2), log(length(EVC_3D_Tree_Ultra$tip.label))),lty = "dashed", lwd = 2, col = "red")

ltt(EVC_3D_MCC_Tree)

EVC_3D_gtt = gtt(EVC_3D_Tree_Ultra)
plot(EVC_3D_gtt)
```

ltt
```{r}
EVC_VP1_LTT = ltt(EVC_VP1_MCC_Tree)
EVC_3D_LTT = ltt(EVC_3D_MCC_Tree)

EVC_VP1_LTT_DF = as.data.frame(cbind(EVC_VP1_LTT[[1]], EVC_VP1_LTT[[2]]))
colnames(EVC_VP1_LTT_DF) <- c("Lineages", "Time")

EVC_3D_LTT_DF = as.data.frame(cbind(EVC_3D_LTT[[1]], EVC_3D_LTT[[2]]))
colnames(EVC_3D_LTT_DF) <- c("Lineages", "Time")

EVC_Delta = max(EVC_VP1_LTT_DF$Time) - max(EVC_3D_LTT_DF$Time)

ggplot() +
  geom_line(aes(x = c(0, max(nodeHeights(EVC_VP1_MCC_Tree))), y = c(log(2), log(length(EVC_VP1_MCC_Tree$tip.label)))), color = "goldenrod", linetype = 2, alpha = 0.5) +
  geom_line(aes(x = c(0 + EVC_Delta, max(nodeHeights(EVC_3D_MCC_Tree)) + EVC_Delta), y = c(log(2), log(length(EVC_3D_MCC_Tree$tip.label)))), color = "lemonchiffon2", linetype = 2, alpha = 0.9) +
  geom_step(data = EVC_VP1_LTT_DF, aes(x = Time, y = log(Lineages)), color = "goldenrod", linewidth = 0.75) +
  geom_step(data = EVC_3D_LTT_DF, aes(x = Time + EVC_Delta, y = log(Lineages)), color = "lemonchiffon2", linewidth = 0.75) +
  ylab("Log Lineages") +
  xlab("Time From Root") +
  theme_prism()
```

gtt
```{r}
EVC_VP1_GTT = gtt(EVC_VP1_MCC_Tree)
EVC_3D_GTT = gtt(EVC_3D_MCC_Tree)

EVC_VP1_GTT_DF = as.data.frame(cbind(EVC_VP1_GTT[[1]], EVC_VP1_GTT[[2]]))
colnames(EVC_VP1_GTT_DF) <- c("Time", "Gamma")

EVC_3D_GTT_DF = as.data.frame(cbind(EVC_3D_GTT[[1]], EVC_3D_GTT[[2]]))
colnames(EVC_3D_GTT_DF) <- c("Time", "Gamma")

EVC_Delta = max(EVC_VP1_GTT_DF$Time) - max(EVC_3D_GTT_DF$Time)

#pdf("EVC_GTT.pdf", height = 3, width = 5)

ggplot() +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "lightgrey") +
  geom_step(data = EVC_VP1_GTT_DF, aes(x = Time, y = Gamma), color = "goldenrod", linewidth = 0.75) +
  geom_step(data = EVC_3D_GTT_DF, aes(x = Time + EVC_Delta, y = Gamma), color = "lemonchiffon2", linewidth = 0.75) +
  ylab("Gamma") +
  xlab("Time From Root") +
  theme_prism()

#dev.off()
```
